|

### 2019-06-28 inst/manuscript/manuscript_supplement.Rmd
Fix atcc analysis in manuscript_supplement.Rmd

# normalise LFQ intensities
atcc_norm <- normalizeCyclicLoess(atcc_filt)

# filter for proteins identified in min 1/3 reps of 10231 EV
# define 'ATCC10231 EV proteins'
atcc1_ev <- filter_na(atcc_norm, op = "<=", 
                      pat = "A10231_EV", val = 2)

# filter for proteins identified in min 1/3 reps of 10231 WCL
# define 'ATCC10231 WCL proteins'
atcc1_w <- filter_na(atcc_norm, op = "<=", 
                     pat = "A10231_W", val = 2)

# filter for proteins identified in min 1/3 reps of 90028 EV
# define 'ATCC90028 EV proteins'
atcc9_ev <- filter_na(atcc_norm, op = "<=", 
                      pat = "A90028_EV", val = 2)

# filter for proteins identified in min 1/3 reps of 90028 WCL
# define 'ATCC90028 WCL proteins'
atcc9_w <- filter_na(atcc_norm, op = "<=", 
                      pat = "A90028_W", val = 2)

# can't impute with so many missing values
# filter for exclusive to 10231 EV or WCL or 90028 EV or WCL (don't impute)
atcc_excl <- filter_na4(atcc_norm, logic = "or", op = "==", 
                        pat1 = "A10231_EV", val1 = 3, 
                        pat2 = "A10231_W", val2 = 3, 
                        pat3 = "A90028_EV", val3 = 3, 
                        pat4 = "A90028_W", val4 = 3)

# filter for proteins with min 1 valid value in all 4 sample types (impute)
atcc_both <- filter_na4(atcc_norm, logic = "and", op = "<=", 
                        pat1 = "A10231_EV", val1 = 2, 
                        pat2 = "A10231_W", val2 = 2, 
                        pat3 = "A90028_EV", val3 = 2, 
                        pat4 = "A90028_W", val4 = 2)

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
atcc_imp <- impute_QRILC(atcc_both)

# recombine imputed proteins and non-imputed proteins in matrix
atcc_de <- rbind(atcc_excl, atcc_imp)
colnames(atcc_de) <- stringi::stri_replace_all_regex(colnames(atcc_de), 
                                                     c("A10231_", "A90028_"), 
                                                     c("A1_", "A9_"), 
                                                     vectorize_all = FALSE)

# see limma user guide section 9.2 for more info about DE
# create design matrix
atcc_samp <- data.frame(T = (rep(c("A1_EV", "A1_WCL", 
                                   "A9_EV", "A9_WCL"), each = 3)))

atcc_design <- stats::model.matrix(~ 0 + T, data = atcc_samp)
colnames(atcc_design) <- c("A1_EV", "A1_W", "A9_EV", "A9_W")

# define sample comparisons of interest
atcc_contrasts <- c("A1_EV - A1_W", "A9_EV - A9_W")

# make all pair-wise comparisons for specified contrasts
# and perform limma::eBayes()
atcc_efit <- limma_eBayes(atcc_de, atcc_design, atcc_contrasts)

# extract overall results
atcc_overall <- get_results(efit = atcc_efit, mat = atcc_de, 
                            p_val = 0.01, lfc = 0, type = "overall")

# extract DE results
a1_res <- get_results(efit = atcc_efit, 
                      mat = atcc_de, 
                      p_val = 0.01, 
                      lfc = 0, 
                      type = "individual")[[1]]

a9_res <- get_results(efit = atcc_efit, 
                      mat = atcc_de, 
                      p_val = 0.01, 
                      lfc = 0, 
                      type = "individual")[[2]]
```

@@

```{r figureS1, fig.cap=f1_cap, fig.width=5.45, fig.height=6}
# construct figure 1
fs1_cap <- "\\textbf{Schematic of the workflow used for the isolation and 
           proteomic analysis of \\textit{C. albicans} EVs and the selection 
           of EV protein markers.} Steps 1 to 6 were performed for each of the 
           four strains examined in this study: DAY286 yeast, ATCC90028 yeast, 
           ATCC10231 yeast, and DAY286 biofilm. The results from the four 
           seperate analyses were combined (step 7) to identify proteins which 
           commonly appeared as EV marker candidates. The MISEV2018 criteria 
           for protein content-based EV characterisation and other 
           considerations were used to select the best \\textit{C. albicans} 
           EV marker proteins."

fs1 <- png::readPNG("figures/figure1.png")

@@

# plot nta distribution for each bio rep
nta_techavg %>% 
  ggplot(aes(x = particle_size, y = Tech_rep_mean, colour = Bio_rep)) + 
  geom_line(size = 0.5) + 
  geom_vline(xintercept = 100, colour = "#999999", linetype = "dashed") +
  facet_wrap(~ Strain, scales = "free", ncol = 2, nrow = 2, 
             labeller = labeller(Strain = figS1_labels)) + 
  theme_classic() + 
  theme(strip.background = element_blank(), 
        legend.position = "bottom", 
        plot.margin = margin(t = 0, l = 0, r = 0.1, b = 0, unit = "inches")) + 
  xlab(label = "Particle size (nm)") + 
  ylab(label = "Particles / mL") + 
  labs(colour = "Biological replicate")
```

@@

# add proper protein functions to table 1
# also add vesicle depleted supernatant info and 
# group proteins according to localisation
wcl_longlist <- wcl_longlist %>% 
  mutate(Function = function_desc2, 
         VDM = c(rep("", 52)), 
         pack_row = c(8, rep(4, 2), 3, rep(2, 2), 7, rep(1, 2), rep(7, 3), 1, 
                      rep(2, 3), 7, 2, 4, rep(2, 2), 7, 3, 2, 5, rep(1, 2), 6, 
                      2, 7, 6, 2, 1, 7, rep(1, 3), 2, 1, 5, 2, 1, 6, 
                      1, 6, 2, 1, 9, 9, rep(1, 3))) %>% 
  arrange(pack_row)
```

@@

# draw table S2
knitr::kable(wcl_longlist[, -10], 
             caption = tabs2_cap, 
             format = "latex", 
             booktabs = TRUE, 
             longtable = TRUE, 
             linesep = "", 
             escape = FALSE) %>% 
  row_spec(row = c(23, 30, 34, 35, 44, 50), underline = TRUE) %>% 
  add_header_above(c(" " = 2, 
                     "log\\\\textsubscript{2}(fold change) EV vs WCL" = 4, 
                     " " = 3), escape = FALSE) %>% 
  kable_styling(font_size = 8, 
                latex_options = c("HOLD_position")) %>% 
  pack_rows("Cytoplasm", 1, 16) %>% 
  pack_rows("Mitochondria", 17, 30) %>% 
  pack_rows("Nucleus", 31, 32) %>% 
  pack_rows("Vacuole", 33, 35) %>% 
  pack_rows("Peroxisome", 36, 37) %>% 
  pack_rows("Proteasome", 38, 41) %>% 
  pack_rows("Cell wall, cell surface, fungal biofilm matrix", 42, 49) %>% 
  pack_rows("Cytoskeleton", 50, 50) %>% 
  pack_rows("Ribosome", 51, 52) %>% 
  footnote(alphabet = c("Protein localisation was inferred from sequence 
                         similarity with \\\\textit{S. cerevisiae} ortholog as 
                         annotated in the Candida Genome Database 
                         (ref:footnote1)", 
                        "Protein localisation was predicted using DeepLoc-1.0 
                         (ref:footnote2)", 
                        "Protein localisation was predicted using BUSCA 
                         (ref:footnote3)"), 
           escape = FALSE, 
           threeparttable = TRUE)
```

>>>>

# normalise LFQ intensities
atcc_norm <- normalizeCyclicLoess(atcc_filt)
```

```{r atcc1, include=FALSE}
# filter for proteins quantified in min 2/3 reps of EV or WCL
a1_norm <- filter_na2(atcc_norm, logic = "or", op = "<=", 
                      pat1 = "A10231_EV", val1 = 1, 
                      pat2 = "A10231_W", val2 = 1)
a1_norm <- a1_norm[, -c(7:12)]

# filter for proteins quantified in min 1/3 reps of EV
# define 'EV proteins'
a1_ev <- filter_na(a1_norm, op = "<=", pat = "EV", val = 2)

# filter for proteins quantified in min 1/3 reps of WCL
# define 'WCL proteins'
a1_wcl <- filter_na(a1_norm, op = "<=", pat = "W", val = 2)

# can't impute with so many missing values
# separate EV/WCL exclusive proteins (don't impute) from
# EV/WCL common proteins (impute)
a1_excl <- filter_na2(a1_norm, logic = "or", op = "==",
                      pat1 = "EV", val1 = 3,
                      pat2 = "W", val2 = 3)

a1_both <- filter_na2(a1_norm, logic = "and", op = "<=",
                      pat1 = "EV", val1 = 2,
                      pat2 = "W", val2 = 2)

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
a1_imp <- impute_QRILC(a1_both)

# recombine imputed proteins and non-imputed proteins in matrix
a1_de <- rbind(a1_excl, a1_imp)
colnames(a1_de) <- gsub("A10231", "A1", colnames(a1_de))

# see limma user guide section 9.2 for more info about DE
# create design matrix
a1_samp <- data.frame(T = (rep(c("EV", "WCL"), each = 3)))

a1_design <- stats::model.matrix(~ 0 + T, data = a1_samp)
colnames(a1_design) <- c("A1_EV", "A1_W")

# define sample comparisons of interest
a1_contrasts <- c("A1_EV - A1_W")

# make all pair-wise comparisons between EV and WCL
# and perform limma::eBayes()
a1_efit <- limma_eBayes(a1_de, design = a1_design, contrasts = a1_contrasts)

# extract DE results
a1_res <- get_results(efit = a1_efit,
                      mat = a1_de,
                      p_val = 0.01,
                      lfc = 0,
                      type = "individual")[[1]]
```

```{r atcc9, include=FALSE}
# filter for proteins quantified in min 2/3 reps of EV or WCL
a9_norm <- filter_na2(atcc_norm, logic = "or", op = "<=",
                      pat1 = "A90028_EV", val1 = 1,
                      pat2 = "A90028_W", val2 = 1)
a9_norm <- a9_norm[, -c(1:6)]

# filter for proteins quantified in min 1/3 reps of EV
# define 'EV proteins'
a9_ev <- filter_na(a9_norm, op = "<=", pat = "EV", val = 2)

# filter for proteins quantified in min 1/3 reps of WCL
# define 'WCL proteins'
a9_wcl <- filter_na(a9_norm, op = "<=", pat = "W", val = 2)

# can't impute with so many missing values
# separate EV/WCL exclusive proteins (don't impute) from
# EV/WCL common proteins (impute)
a9_excl <- filter_na2(a9_norm, logic = "or", op = "==",
                      pat1 = "EV", val1 = 3,
                      pat2 = "W", val2 = 3)

a9_both <- filter_na2(a9_norm, logic = "and", op = "<=",
                      pat1 = "EV", val1 = 2,
                      pat2 = "W", val2 = 2)

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
a9_imp <- impute_QRILC(a9_both)

# recombine imputed proteins and non-imputed proteins in matrix
a9_de <- rbind(a9_excl, a9_imp)
colnames(a9_de) <- gsub("A90028", "A9", colnames(a9_de))

# see limma user guide section 9.2 for more info about DE
# create design matrix
a9_samp <- data.frame(T = (rep(c("EV", "WCL"), each = 3)))

a9_design <- stats::model.matrix(~ 0 + T, data = a9_samp)
colnames(a9_design) <- c("A9_EV", "A9_W")

# define sample comparisons of interest
a9_contrasts <- c("A9_EV - A9_W")

# make all pair-wise comparisons between EV and WCL
# and perform limma::eBayes()
a9_efit <- limma_eBayes(a9_de, design = a9_design, contrasts = a9_contrasts)

# extract DE results
a9_res <- get_results(efit = a9_efit,
                      mat = a9_de,
                      p_val = 0.01,
                      lfc = 0,
                      type = "individual")[[1]]
```

@@

```{r figureS1, fig.cap=fs1_cap, fig.width=5.45, fig.height=6}
# construct figure 1
fs1_cap <- "\\textbf{Schematic of the workflow used for the isolation and 
           proteomic analysis of \\textit{C. albicans} EVs and the selection 
           of EV protein markers.} Steps 1 to 6 were performed for each of the 
           four strains examined in this study: DAY286 yeast, ATCC90028 yeast, 
           ATCC10231 yeast, and DAY286 biofilm. The results from the four 
           seperate analyses were combined (step 7) to identify proteins which 
           commonly appeared as EV marker candidates. The MISEV2018 criteria 
           for protein content-based EV characterisation and other 
           considerations were used to select the best \\textit{C. albicans} 
           EV marker proteins."

fs1 <- png::readPNG("figures/figures1.png")

@@

# plot nta distribution for each bio rep
nta_techavg %>% 
  ggplot(aes(x = particle_size, y = Tech_rep_mean, colour = Bio_rep)) + 
  geom_line(size = 0.5) + 
  geom_vline(xintercept = 100, colour = "#999999", linetype = "dashed") +
  facet_wrap(~ Strain, scales = "free", ncol = 2, nrow = 2, 
             labeller = labeller(Strain = figS2_labels)) + 
  theme_classic() + 
  theme(strip.background = element_blank(), 
        legend.position = "bottom", 
        plot.margin = margin(t = 0, l = 0, r = 0.1, b = 0, unit = "inches")) + 
  xlab(label = "Particle size (nm)") + 
  ylab(label = "Particles / mL") + 
  labs(colour = "Biological replicate")
```

@@

# add proper protein functions to table 1
# also add vesicle depleted supernatant info and 
# group proteins according to localisation
wcl_longlist2 <- wcl_longlist %>% 
  mutate(Function = function_desc2, 
         VDM = c(rep("", 40)), 
         pack_row = c(rep(4, 2), 3, rep(2, 2), 7, rep(1, 2), rep(7, 2), 1, 
                      rep(2, 4), 7, 3, 2, 5, 1, 
                      2, 6, 2, 1, 7, rep(1, 2), 2, 1, 5, 1, 
                      1, 6, 2, 1, rep(8, 2), rep(1, 3))) %>% 
  arrange(pack_row)
```

@@

# draw table S2
knitr::kable(wcl_longlist2[, -10], 
             caption = tabs2_cap, 
             format = "latex", 
             booktabs = TRUE, 
             longtable = TRUE, 
             linesep = "", 
             escape = FALSE) %>% 
  row_spec(row = c(20, 25, 29, 36), underline = TRUE) %>% 
  add_header_above(c(" " = 2, 
                     "log\\\\textsubscript{2}(fold change) EV vs WCL" = 4, 
                     " " = 3), escape = FALSE) %>% 
  kable_styling(font_size = 8, 
                latex_options = c("HOLD_position")) %>% 
  pack_rows("Cytoplasm", 1, 14) %>% 
  pack_rows("Mitochondria", 15, 25) %>% 
  pack_rows("Nucleus", 26, 27) %>% 
  pack_rows("Vacuole", 28, 29) %>% 
  pack_rows("Peroxisome", 30, 31) %>% 
  pack_rows("Proteasome", 32, 33) %>% 
  pack_rows("Cell wall, cell surface, fungal biofilm matrix", 34, 38) %>% 
  pack_rows("Ribosome", 39, 40) %>% 
  footnote(alphabet = c("Protein localisation was inferred from sequence 
                         similarity with \\\\textit{S. cerevisiae} ortholog as 
                         annotated in the Candida Genome Database 
                         (ref:footnote1)", 
                        "Protein localisation was predicted using DeepLoc-1.0 
                         (ref:footnote2)", 
                        "Protein localisation was predicted using BUSCA 
                         (ref:footnote3)"), 
           escape = FALSE, 
           threeparttable = TRUE)
```
